---
title: "Class 9 - Structural Bioinformatics"
author: "Jenny"
format: html
---

## THe RCSB Protein Data Bank  (PDB )

Protein structures by X-ray crystalography dominante this database. We are going to use Molstar
Skipping questions 1-3 since the website was too slow for us

* Computer doesn't allow me to open WebGL since my browser seems to be outdated so I can't do this step...
Without having access to this website I'm not able to answer questions 4-7

## 3. Introduction to Bio3D in R

Bio3D is an R package for structural bioinformaics. To use it we need to call it up with the 'library()' function ( just like any package)

```{r}
library("bio3d")
```

```{r}
pdb <- read.pdb("1hsg")
pdb
```

> Q7. How many amino acid residues are there in this pdb object ? 
          198
          
> Q8. Name one of the two non-protein residues ?
          mk1

>Q9. How many protein chains are in this structure ?
          2

```{r}
attributes(pdb)
```

```{r}
head(pdb$atom)
```

## Comparative structure analysis of Adenylate Kinase
    Adenylate kinase structure (PDB code: 1AKE)

Make sure to install the packages first

>Q10. Which of the packages above is found only on BioConductor and not CRAN ?
          msa

>Q11. Which of the above packages is not found on BioConductor or CRAN ?
          none
          
>Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket ?
          TRUE

First we get it's primary sequence :
```{r}
library(bio3d)
aa <- get.seq("1ake_A")
```
```{r}
aa
```

>Q13.How many amino acids are in this sequence, i.e. how long is this sequence ?
      214
      
```{r}
# Blast or hammer search
b <- blast.pdb(aa)
```

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

```{r}
# List out some 'top hits'
head(hits$pdb.id)
```



```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile = "msa")
```

```{r}
ids <- basename.pdb(pdbs$id)
plot(pdbs, labels=ids)
par(mar=c(1,1,1,1))
```

```{r}
anno <- pdb.annotate(ids)
unique(anno$source)
```

## PCA

```{r}
pc.xray <- pca(pdbs)
plot(pc.xray)
```

```{r}
rd <- rmsd(pdbs)
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

```{r}
library(ggplot2)
library(ggrepel)

df <- data.frame(PC1=pc.xray$z[,1], 
                 PC2=pc.xray$z[,2], 
                 col=as.factor(grps.rd),
                 ids=ids)

p <- ggplot(df) + 
  aes(PC1, PC2, col=col, label=ids) +
  geom_point(size=2) +
  geom_text_repel(max.overlaps = 20) +
  theme(legend.position = "none")
p
```


