---
title: "Halloween Candy Project"
author: "Jenny"
format: pdf
---

## Import candy data

```{r}
candy_file <- "candy-data.csv"
candy <-read.csv(candy_file,row.names=1)
head(candy)
```

>Q1. How many different candy types are there in the data set ?

```{r}
nrow(candy)
```

>Q2. How many fruity candy types are in the dataset ?

```{r}
sum(candy$fruity)
```

```{r}
candy["Twix", ]$winpercent
```

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value ?

```{r}
candy["WelchÕs Fruit Snacks",]$winpercent
```

>Q4. What is the winpercent value for “Kit Kat” ?

```{r}
candy["Kit Kat",]$winpercent
```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars” ?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```

```{r}
#install.packages("skimr")
```

```{r}
library("skimr")
skim(candy)
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset ?
        It would be winpercent since the numbers for this column differ from that of the rest of the columns
        
>Q7. What do you think a zero and one represent for the candy$chocolate column ?
        Zero represents FALSE and one would represent TRUE

```{r}
skim(candy$chocolate)
```

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

```{r}
library(ggplot2)

ggplot(candy) + aes(winpercent) + geom_histogram(bins=10)
```

>Q9. Is the distribution of winpercent values symmetrical ?
            No, the distribution of winpercent values is not symmetrical

>Q10. Is the center of the distribution above or below 50% ?
            The center of the distribution is above 50%

>Q11. On average is chocolate candy higher or lower ranked than fruity candy ?
          Chocolate candy is higher ranked than fruity candy

```{r}
chocolate.inds <- as.logical(candy$chocolate)
chocolate.wins <- candy[chocolate.inds,]$winpercent
```
```{r}
mean(chocolate.wins)
```

```{r}
fruity.inds <-as.logical(candy$fruity)
candy.wins <- candy[fruity.inds,]$winpercent
```
```{r}
mean(candy.wins)
```

>Q12. Is this difference statistically significant ?
        Yes

```{r}
t.test(chocolate.wins, candy.wins)
```

## Overall Candy Rankings

>Q13. What are the five least liked candy types in this set ?

```{r}
tail(candy[order(candy$winpercent),], n=5)
```

>Q14. What are the top 5 all time favorite candy types out of this set ?

```{r}
library(dplyr)
candy %>% 
arrange(winpercent) %>%
head(5)
```

>Q15. a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent ?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col() 
```

First setup some colors for differen candy types :

```{r}
my_cols=rep("black",nrow(candy))

my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

my_cols
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

>Q17. What is the worst ranked chocolate candy ?
        The worst ranked chocolate candy was Sixlets

>Q18. What is the best ranked fruity candy ?
        The best ranked fruity candy was Starburst

```{r}
library(ggrepel)
```

```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck ?
      The most bang for your buck would be Reeses Miniatures

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular ?
        5 most expensives candies are Nik L Nip, Nestle Smarties, Ring Pop, HersheyÕs Krackel and HersheyÕs Milk Chocolate. The least popular from these is Nik L Nip 
          
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

## Exploring the correlation structure

```{r}
#install.packages("corrplot")
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values) ?
            The two variables that are anti-correlated are fruity & chocolate

>Q23. Similarly, what two variables are most positively correlated ?
            The two variables are most positively correlated are chocolate & chocolate; basically if the variable is the same, then it will give you a positive correlation
            
## Principal Component Analysis

```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2])
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
#install.packages("plotly")
#library(plotly)
```

```{r}
#ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

>Q24.What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?
          The original variable that are picked up strongly by PC1 in the positive direction are fruity, hard and pluribus. It does not make sense to me since these variables weren't the ones that were  high as shown in the column plot from above.

